---
title: "Palmer Penguins: An R Analysis"
author: "Zaakeer Ahmad M"
date: 2026-09-28
---

## Question and data

This post examines how body mass and bill length vary across the three species
in the Palmer Penguins dataset. The data come from the
[Palmer Penguins project](https://allisonhorst.github.io/palmerpenguins/) and
were collected by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER.
The dataset is released under the
[CC0 1.0 public-domain dedication](https://creativecommons.org/publicdomain/zero/1.0/).
It is included in the `palmerpenguins` R package, so rendering this post does
not download a separate data file.

## Prepare the data


```{r}
#| label: setup-r
#| message: false
#| warning: false

library(palmerpenguins)
library(tidyverse)
```

Removing rows with missing values

```{r}
#| label: clean-penguins-r

analysis_penguins <- penguins |>
  drop_na(species, bill_length_mm, body_mass_g, sex)

tibble(
  original_rows = nrow(penguins),
  complete_rows = nrow(analysis_penguins)
)
```

## Compare the species

Then we calculate the sample size and mean body measurements for each
species.

```{r}
#| label: tbl-species-summary-r
#| tbl-cap: "Sample size and mean body measurements by penguin species."

species_summary <- analysis_penguins |>
  group_by(species) |>
  summarise(
    n = n(),
    mean_body_mass_g = mean(body_mass_g),
    mean_bill_length_mm = mean(bill_length_mm),
    .groups = "drop"
  ) |>
  mutate(across(starts_with("mean_"), ~ round(.x, 1)))

species_summary
```

Gentoo penguins are substantially heavier on average: approximately 5,092 g,
compared with 3,706 g for Adelie and 3,733 g for Chinstrap penguins. Chinstrap
penguins have the longest mean bill at approximately 48.8 mm.

The figure compares bill length with body mass. Colour separates the species,
making both the between-species differences and the within-species trends
visible.

```{r}
#| label: fig-r-bill-mass
#| fig-cap: "Bill length and body mass for Palmer Penguins, coloured by species."
#| warning: false

ggplot(
  analysis_penguins,
  aes(x = bill_length_mm, y = body_mass_g, colour = species)
) +
  geom_point(alpha = 0.75, size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Bill length (mm)",
    y = "Body mass (g)",
    colour = "Species"
  ) +
  theme_minimal(base_size = 12)
```

To quantify those within-species trends, I calculate the correlation between
bill length and body mass separately for each species.

```{r}
#| label: tbl-species-correlations-r
#| tbl-cap: "Correlation between bill length and body mass within each species."

species_correlations <- analysis_penguins |>
  group_by(species) |>
  summarise(
    correlation = cor(bill_length_mm, body_mass_g),
    .groups = "drop"
  ) |>
  mutate(correlation = round(correlation, 3))

species_correlations
```

Bill length and body mass are positively corelated within all three species.
The relationship is strongest for Gentoo penguins (correlation 0.667), followed
by Adelie (0.544) and Chinstrap penguins (0.514). Overall, species explains a
difference in typical body mass, while bill length still provides useful
information about how it is corealted with body mass within each species.
