---
title: "Palmer Penguins with R and Python"
author: "Zaakeer Ahmad M"
date: 2026-09-28
---

## Passing an object between languages

This bonus post uses R and Python in the same Quarto document. 

The data come from the
[Palmer Penguins project](https://allisonhorst.github.io/palmerpenguins/) and
are available under the
[CC0 1.0 public-domain dedication](https://creativecommons.org/publicdomain/zero/1.0/).

```{r}
#| label: setup-mixed-analysis
#| message: false
#| warning: false

library(dplyr)
library(palmerpenguins)
library(reticulate)

project_python <- Sys.which("python")
use_python(project_python, required = TRUE)
```

## Step 1: Compute a summary in R

R calculates the mean body mass for each species. The resulting
`species_mass_r` object will be made available to Python by `reticulate`.

```{r}
#| label: tbl-mass-summary-r
#| tbl-cap: "Mean penguin body mass calculated in R."

species_mass_r <- penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species) |>
  summarise(
    mean_body_mass_g = round(mean(body_mass_g), 1),
    .groups = "drop"
  )

species_mass_r
```

## Step 2: Use the R object in Python

In a Python chunk, `r.species_mass_r` retrieves the object created by R.
Python uses it to express every mean relative to the lightest species and
stores the result as `species_mass_python`.

```{python}
#| label: tbl-relative-mass-python
#| tbl-cap: "Relative mean body mass calculated in Python from the R object."

species_mass_python = r.species_mass_r.copy()
lightest_mean = species_mass_python["mean_body_mass_g"].min()
species_mass_python["relative_to_lightest"] = (
    species_mass_python["mean_body_mass_g"] / lightest_mean
).round(2)

species_mass_python
```

## Step 3: Return the Python result to R

The `py$species_mass_python` expression brings the Python result back into R.
This confirms that the same object moved from R to Python and then back to R.

```{r}
#| label: tbl-python-result-in-r
#| tbl-cap: "The Python result read back into R."

result_from_python <- py$species_mass_python
result_from_python
```

Gentoo penguins have the largest mean body mass. Their mean is approximately
1.37 times the mean for Adelie penguins, the lightest species in this group.
This result was possible because the R summary was passed directly to Python
rather than being written to and reread from a separate file.
